---
title: "Final Project"
output: 
  flexdashboard::flex_dashboard:
    theme: cerulean
    orientation: columns
    vertical_layout: fill
    social: ["twitter", "facebook"]
    source_code: https://github.com/aashayadav/flexdashboard/blob/master/flexdashboard.Rmd
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(rio)
library(here)
library(naniar)
library(colorblindr)
library(tigris)
library(geofacet)
library(cowplot)
library(ggrepel)
library(DT)
# with plotly you can use ggplotly(put name of plot). It will add some interactivity
# package (DT) and package (reactable): these are for making tables in your dashboard
# you can use select first then just pipe it to reactable()
NCSH <- import(here("data","2017-2018 NSCH_Topical_DRC_Dec 2019.csv"))
state <- import(here("data","fips_state.csv"))
final_data <- left_join(NCSH, state, by = "FIPSST")
# Data under 5
data_under5 <- final_data %>%
  select(State, SC_CSHCN, ACE2more_1718, SC_AGE_YEARS) %>%
  filter(SC_AGE_YEARS <= 5)
# Numeric to character

data_under5$SC_CSHCN[which(data_under5$SC_CSHCN==1)]<- "Sp Health Needs"
data_under5$SC_CSHCN[which(data_under5$SC_CSHCN==2)]<-"No Sp Health needs"

data_under5$ACE2more_1718[which(data_under5$ACE2more_1718==1)]<-"Exp No ACE"
data_under5$ACE2more_1718[which(data_under5$ACE2more_1718==2)]<-"Exp 1 ACE"
data_under5$ACE2more_1718[which(data_under5$ACE2more_1718==3)]<-"Exp 2 ACE"
#missing data
data_under5 <- data_under5 %>%
  replace_with_na(replace = list(ACE2more_1718 = 99)) %>%
  drop_na()
# count adverse childhood experiences

data_under5 <- data_under5 %>%
  select(State, SC_CSHCN, ACE2more_1718) %>%
  group_by(State, SC_CSHCN) %>%
  count(ACE2more_1718)
# percentage

data_under5 <- data_under5 %>%
  mutate(Percentage = (n/sum(n))*100)


```
Sidebar {.sidebar}
=====================================
The National Survey of Children's Health (NSCH) provides data on multiple, intersecting aspects of children's lives including mental and physical health. This dashboard presents visualization of Adverse Childhood Experiences (ACEs) in Children with Special Health Needs(CSHN) from 0-5 years old, using publicly avaialble NSCH dataset. Dataset can be accessed at their [website](https://www.childhealthdata.org/learn-about-the-nsch/NSCH). This dashboard presents three visualisations: Nationwide prevalence of ACEs in CSHN, comparison between East and West Coast, and comparison between 2016 and 2017.
For more information on ACEs in Children with Special Health Needs, check [Childtrend](https://www.childtrends.org/children-with-special-health-care-needs-are-more-likely-to-have-adverse-childhood-experiences). Dashboard will be shared with colleagues in Early Intervention at the University of Oregon.
Before finalizing the plots different versions were tried using differnt plots, colors, title and legend position, theme etc.


Nationwide {data-icon="fa-map"} 
=====================================  

Row
-------------------------------------
   
### State wise display of ACE and Special Health Needs



```{r, fig.height=15, fig.width=20}

ggplot(data_under5, aes(x = ACE2more_1718,
                        y = Percentage,
                        fill = SC_CSHCN)) +
  geom_col(position = "dodge") +
  coord_flip() +
  theme_bw() +
  facet_geo(~State) +
  theme_bw() +
  theme(strip.text.x = element_text(size=15),
        strip.background = element_rect(color = "black"),
        axis.title.y = element_blank(),
        text = element_text(size = 20),
        legend.title = element_blank()) +
  labs(title = "Children under 5 with special health needs likely to experience higher adverse childhood experiences",
       caption = "Data Source: NSCH 2017-18 data") +
  scale_fill_OkabeIto()
  
```

East and West Coast {data-icon="fa-balance-scale"} 
=====================================  

Column {.tabset}
-------------------------------------

### East and West Coast

```{r}
data_under5$region <- recode(data_under5$State,
                              "California" = "West Coast",
                              "Oregon" = "West Coast",
                              "Alaska" = "West Coast",
                              "Washington" = "West Coast")

p1 <- data_under5 %>% 
  group_by(State) %>% 
  filter(region == "West Coast") %>% 
  ungroup() %>% 
ggplot(aes(fct_reorder(State, Percentage), Percentage)) +
  geom_col(aes(fill = ACE2more_1718), position = "dodge") +
  coord_flip() +
  facet_wrap(~SC_CSHCN) +
  theme(strip.text.x = element_text(size=8),
        strip.background = element_rect(color = "black"),
        legend.position = "none",
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        text = element_text(size = 8),
        legend.title = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()) +
  theme(plot.title = element_text(hjust = 0.50)) +
  scale_fill_brewer(palette = "Set2")

data_under5$region <- recode(data_under5$State,
                             "New York" = "East Coast",
                             "New Jersey" = "East Coast",
                             "Massachusetts" = "East Coast",
                             "Pennsylvania" = "East Coast")

p2 <- data_under5 %>% 
  group_by(State) %>% 
  filter(region == "East Coast") %>% 
  ungroup() %>% 
ggplot(aes(fct_reorder(State, Percentage), Percentage)) +
  geom_col(aes(fill = ACE2more_1718), position = "dodge") +
  coord_flip() +
  facet_wrap(~SC_CSHCN) +
  theme(strip.text.x = element_text(size=8),
        strip.background = element_rect(color = "black"),
        legend.position = "bottom",
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        text = element_text(size = 8),
        legend.title = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()) +
  theme(plot.title = element_text(hjust = 0.10)) +
  scale_fill_brewer(palette = "Set2")

plot_row <- plot_grid(p1, p2, ncol=1, labels = c())

title <- ggdraw() + 
  draw_label(
    "Children under 5 with special health needs likely to experience 1\n or more adverse\n childhood experiences in East and West Coast\n(Data Source: NSCH 2016 & NSCH 2017)",
    x = 0, hjust = 0) + 
  theme(
    plot.margin = margin(0, 0, 0, 7)
    )
  
plot_grid(
  title, plot_row,
  ncol = 1,
  rel_heights = c(0.1, 1)
  )
```

2016-2017 {data-icon="fa-calendar"}
=====================================   

### 2016-2017 Comparison

```{r, fig.width=12, fig.height=10}

NCSH_17_p2 <- NCSH%>%
  select(FIPSST, SC_CSHCN, ACE2more_1718, SC_AGE_YEARS, ACEdivorce_1718, ACEdeath_1718, ACEjail_1718, ACEdomviol_1718, ACEneighviol_1718, ACEmhealth_1718, ACEdrug_1718 )%>%
  filter(SC_AGE_YEARS <= 5, SC_CSHCN == 1, ACE2more_1718 == 1 ) %>%
  rename(SC_CSHCN_1718 = "SC_CSHCN")

NCSH_17_p2 <- left_join(NCSH_17_p2, state, by = "FIPSST")

NCSH_17_p2 <- NCSH_17_p2 %>%
  select(State, ACE2more_1718) %>%
  group_by(State) %>%
  count()%>%
  rename(ACE_17 = "n")

NCSH_16_p2 <- import(here("data","2016NSCHTopical_DRCv3_Sep2018.csv")) %>%
  select(FIPSST, SC_CSHCN, ACE2more_16, SC_AGE_YEARS, ACEdivorce_16, ACEdeath_16, ACEjail_16, ACEdomviol_16, ACEneighviol_16, ACEmhealth_16, ACEdrug_16) %>%
  filter(SC_AGE_YEARS <= 5, SC_CSHCN == 1, ACE2more_16 ==1) %>%
  rename(SC_CSHCN_16 = SC_CSHCN)

NCSH_16_p2 <- left_join(NCSH_16_p2, state, by = "FIPSST") %>%
  select(State, ACE2more_16) %>%
  group_by(State) %>%
  count() %>%
  rename(ACE_16 = "n")

data_1618 <- full_join(NCSH_17_p2, NCSH_16_p2, by = "State") %>%
  filter(State%in%c("Alaska", "Oregon", "Washington", "California", "New York", "New Jersey", "Massachusetts", "Pennsylvania"))


library(scales)
theme_set(theme_classic())

colnames(data_1618) <- c("State", "ACE_16", "ACE_17")
left_label <- paste(data_1618$State, round(data_1618$`ACE_16`),sep=", ")
right_label <- paste(data_1618$State, round(data_1618$`ACE_17`),sep=", ")
data_1618$class <- ifelse((data_1618$`ACE_17` - data_1618$`ACE_16`) < 0, "red", "green")

# Plot

p2 <- ggplot(data_1618) + geom_segment(aes(x=1, xend=2, y=`ACE_16`, yend=`ACE_17`,
                                          col=class), size= 1, show.legend=F) + 
  geom_vline(xintercept=1, linetype="dashed", size=.5) +
  geom_vline(xintercept=2, linetype="dashed", size=.5) +
  scale_color_manual(labels = c("Up", "Down"),
                     values = c("green"="#00ba38", "red"="#f8766d")) +
  labs(x="", y="Number of children", 
       title = "Number of reported Sp Health needs children with Adverse Childhood Experiences (birth to 5) in\nEast Coast and West Coast\n") +
  xlim(.5, 2.5) + ylim(0,(1.1*(max(data_1618$`ACE_16`, data_1618$`ACE_17`))))

# Add texts
p2 <- p2 + geom_text(label=left_label, y=data_1618$`ACE_16`, x=rep(1, NROW(df)), 
                     hjust=1.1, size=5)
p2 <- p2 + geom_text_repel(label=right_label, y=data_1618$`ACE_17`, x=rep(2, NROW(df)),
                     hjust=-0.1, size=5)
p2 <- p2 + geom_text(label="TIME 1 (2016)", x=1, y=1.1*(max(data_1618$`ACE_16`,
                                                     data_1618$`ACE_17`)), 
                     hjust=1.2, size=5)
p2 <- p2 + geom_text(label="TIME 2 (2017)", x=2, y=1.1*(max(data_1618$`ACE_16`,
                                                     data_1618$`ACE_17`)), 
                     hjust=-0.1, size=5)  

p2 + theme(panel.background = element_blank(), 
           panel.grid = element_blank(),
           axis.ticks = element_blank(),
           axis.text.x = element_blank(),
           panel.border = element_blank(),
           axis.text.y = element_text(size = 20),
           axis.title = element_text(size = 10),
           plot.title = element_text(color = "black", size = 15, 
                                     face = "bold", hjust = 0.5 ),
           plot.margin = unit(c(2,4,2,4), "cm"))
```

Versions {data-icon="fa-list"}
=====================================   
Column {data-width=40%}
------------------------------------
### Version 1: default color, no title

```{r}
data_under5$region <- recode(data_under5$State,
                              "California" = "West Coast",
                              "Oregon" = "West Coast",
                              "Alaska" = "West Coast",
                              "Washington" = "West Coast")

data_under5 %>% 
  group_by(State) %>% 
  filter(region == "West Coast") %>% 
  ungroup() %>% 
ggplot(aes(fct_reorder(State, Percentage), Percentage)) +
  geom_col(aes(fill = ACE2more_1718), position = "dodge") +
  coord_flip() +
  facet_wrap(~SC_CSHCN) +
  theme(strip.text.x = element_text(size=8),
        strip.background = element_rect(color = "black"),
        legend.position = "bottom",
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        text = element_text(size = 8),
        legend.title = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()) +
  theme(plot.title = element_text(hjust = 0.50))
  
```

### Version 2: With title and caption
```{r}
data_under5$region <- recode(data_under5$State,
                              "California" = "West Coast",
                              "Oregon" = "West Coast",
                              "Alaska" = "West Coast",
                              "Washington" = "West Coast")

data_under5 %>% 
  group_by(State) %>% 
  filter(region == "West Coast") %>% 
  ungroup() %>% 
ggplot(aes(fct_reorder(State, Percentage), Percentage)) +
  geom_col(aes(fill = ACE2more_1718), position = "dodge") +
  coord_flip() +
  facet_wrap(~SC_CSHCN) +
  theme(strip.text.x = element_text(size=8),
        strip.background = element_rect(color = "black"),
        legend.position = "bottom",
        axis.title.x = element_blank(),
        axis.title.y = element_blank(),
        text = element_text(size = 8),
        legend.title = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()) +
  theme(plot.title = element_text(hjust = 0.50)) +
  labs(title = "Children under 5 with special health needs likely to experience 1 or more adverse\n childhood experiences in East and West Coast (Data Source: NSCH 2016 & NSCH 2017)",
       caption = "Data Source: NSCH 2017-18")
  
```

Column {data-width=60%}
----------------------------------------
### Version 3: State wise dispay with facet_wrap {data-height=100%}

```{r}

ggplot(data_under5, aes(x = ACE2more_1718,
                        y = Percentage,
                        fill = SC_CSHCN)) +
  geom_col(position = "dodge") +
  coord_flip() +
  theme_bw() +
  facet_wrap(~State) +
  theme_bw() +
  theme(strip.text.x = element_text(size=7),
        strip.background = element_rect(color = "black"),
        axis.title.y = element_blank(),
        text = element_text(size = 10),
        legend.title = element_blank()) +
  labs(title = "Children under 5 with special health needs\nlikely to experience higher adverse childhood experiences",
       caption = "Data Source: NSCH 2017-18 data") 
```

Data Table {data-icon="fa-database"}
=====================================   

```{r}
datatable(data_under5,
          caption = "NSCH Data:Birth to 5-year-old",
          rownames = T,
          filter = "top",
          options = list(pageLenth = 10))
```



